---
title: "likert Package Examples"
author: "Jake Chanenson"
date: "5/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(foreign)      
library(ggplot2)      
library(colorspace)   #better colors for ggplot
library(tidyverse)    
library(dplyr)
library(RColorBrewer)
library(grid)
library(likert) # for creating a likert plot
```


## Usuing Long data
```{r}
load(file = "../data/sample-likert-data.rda") #our data, see likert-data-generation.Rmd for more info
head(dental_hyg)

## Rename Cols
dental_hyg  <- dental_hyg  %>% 
  rename("How Often Respondents Use Mouth Wash" = mouth_wash,
          "How Often Respondents Brush Their Teeth" = tooth_brush,
          "How Often Respondents Use Floss" = floss)

## New df for not grouped
dh <- dental_hyg %>% select(-age)

## Not Grouped
plot(likert(dh),legend.position="right")
ggsave("../imgs/long_data.png", height = 3, width = 9)

## Grouped
plot(likert(dental_hyg[,c(1:3)], grouping = dental_hyg[,4] ), legend.position="right")
ggsave("../imgs/long_data_grouped.png", height = 3, width = 9)
```

## Using Pre-Surmized Data
```{r}
## Load df
load(file = "../data/sample-likert-data.rda") #our data, see likert-data-generation.Rmd for more info
head(cookie_data)

plot(likert(summary = cookie_data), plot.percent.neutral=FALSE, legend.position="right")
ggsave("../imgs/ps_data.png", height = 3, width = 9)
```


## Modifying The Image

### Changing The Color Scheme and Color Ordering On The Graph
```{r Better-Colors, message=F}
load(file = "../data/sample-likert-data.rda") #our data, see likert-data-generation.Rmd for more info

## Rename Cols
dental_hyg  <- dental_hyg  %>% 
  rename("How Often Respondents Use Mouth Wash" = mouth_wash,
          "How Often Respondents Brush Their Teeth" = tooth_brush,
          "How Often Respondents Use Floss" = floss)

## New df for not grouped
dh <- dental_hyg %>% select(-age)

## Blue-Never, Red-Always
plot(likert(dh), legend.position="right") +
  scale_fill_manual(values = brewer.pal(n=5,"RdYlBu"), breaks = c( "Always", "Often", "Sometimes","Rarely", "Never")) #order and color the likert boxes
ggsave("../imgs/color_b-r.png", height = 3, width = 9)

## Red-Never, Blue-Always  
plot(likert(dh), legend.position="right") +
  scale_fill_manual(values = brewer.pal(n=5,"RdYlBu"), breaks = c("Never", "Rarely", "Sometimes", "Often", "Always")) #order and color the likert boxes
ggsave("../imgs/color_r-b.png", height = 3, width = 9)
```

### Fix Legend
```{r Legend-Reverse, message=F}
## Red-Never, Blue-Always  
plot(likert(dh), legend.position="right") +
  scale_fill_manual(values = brewer.pal(n=5,"RdYlBu"), breaks = c("Never", "Rarely", "Sometimes", "Often", "Always")) +
  guides(fill = guide_legend(title="Responses", reverse = FALSE)) #to reverse the order in the legend
ggsave("../imgs/guide_r-b.png", height = 3, width = 9)

## Red-Never, Blue-Always  
plot(likert(dh), legend.position="right") +
  scale_fill_manual(values = brewer.pal(n=5,"RdYlBu"), breaks = c("Never", "Rarely", "Sometimes", "Often", "Always")) +
  guides(fill = guide_legend(title="Responses", reverse = TRUE)) #to reverse the order in the legend
ggsave("../imgs/guide_b-r.png", height = 3, width = 9)
```

### Reversing The Levels On The Graph
```{r reverse-levels, message=F}
## Reverse Levels
dh.reverse <- reverse.levels(dh)

## Red-Never, Blue-Always reversed
plot(likert(dh.reverse), legend.position="right") +
  scale_fill_manual(values = brewer.pal(n=5,"RdYlBu"), breaks = c("Never", "Rarely", "Sometimes", "Often", "Always")) +
  guides(fill = guide_legend(title="Responses", reverse = TRUE)) #to reverse the order in the legend
ggsave("../imgs/reverse_wrong.png", height = 3, width = 9)
```


### Changing The Ordering Of The Categorical Variables (Not Grouped)
```{r reorder-not-group, message=F}
## Reverse Levels
dh.reverse <- reverse.levels(dh)

plot(likert(dh.reverse), group.order=names(dh), legend.position="right") + #group.order changes the order of the variables
  scale_fill_manual(values = brewer.pal(n=5,"RdYlBu"), breaks = c("Never", "Rarely", "Sometimes", "Often", "Always")) +
  guides(fill = guide_legend(title="Responses", reverse = TRUE)) 
ggsave("../imgs/reverse_right.png", height = 3, width = 9)
```

### Changing The Ordering Of The Categorical Variables (Grouped)
```{r reorder-grouped, message=F}
## Reorder Levels
dental_hyg.reorder <- dental_hyg[,c(3,2,1,4)]

## Normal Grouped Plot
plot(likert(dental_hyg[,c(1:3)], grouping = dental_hyg[,4]), plot.percent.low = FALSE, plot.percent.high = FALSE, plot.percent.neutral = FALSE, legend.position="right") +
  scale_fill_manual(values = brewer.pal(n=5,"RdYlBu"), breaks = c("Never", "Rarely", "Sometimes", "Often", "Always")) +
  guides(fill = guide_legend(title="Responses", reverse = TRUE))
ggsave("../imgs/reorder_g1.png", height = 3, width = 9)

## Reorderd Group Plot
plot(likert(dental_hyg.reorder[,c(1:3)], grouping = dental_hyg.reorder[,4]), plot.percent.low = FALSE, plot.percent.high = FALSE, plot.percent.neutral = FALSE, legend.position="right") +
  scale_fill_manual(values = brewer.pal(n=5,"RdYlBu"), breaks = c("Never", "Rarely", "Sometimes", "Often", "Always")) +
  guides(fill = guide_legend(title="Responses", reverse = TRUE))
ggsave("../imgs/reorder_g2.png", height = 3, width = 9)
```



### Changing The Ordering Of Grouping Within Facets (grouped data only)
```{r Grouping-Within-Facets, message=F}
## Reverse Levels
dental_hyg.reverse <- reverse.levels(dental_hyg)

plot(likert(dental_hyg[,c(1:3)], grouping = dental_hyg.reverse[,4]), plot.percent.low = FALSE, plot.percent.high = FALSE, plot.percent.neutral = FALSE, legend.position="right") + #note grouping = dental_hyg.reverse
  scale_fill_manual(values = brewer.pal(n=5,"RdYlBu"), breaks = c("Never", "Rarely", "Sometimes", "Often", "Always")) +
  guides(fill = guide_legend(title="Responses", reverse = TRUE))
ggsave("../imgs/reorder_g3.png", height = 3, width = 9)                            
```

### No center on diverge and 100% complete
```{r no-center, message=F}
## No Center
plot(likert(dental_hyg[,c(1:3)], grouping = dental_hyg[,4]), plot.percent.low = FALSE, plot.percent.high = FALSE, plot.percent.neutral = FALSE, include.center=FALSE, legend.position="right") +
  scale_fill_manual(values = brewer.pal(n=5,"RdYlBu"), breaks = c("Never", "Rarely", "Sometimes", "Often", "Always")) +
  guides(fill = guide_legend(title="Responses", reverse = TRUE))
ggsave("../imgs/no_center_diverge.png", height = 3, width = 9)

## 100% Complete
plot(likert(dental_hyg[,c(1:3)], grouping = dental_hyg[,4]), plot.percent.low = FALSE, plot.percent.high = FALSE, plot.percent.neutral = FALSE, centered=FALSE, legend.position="right") +
  scale_fill_manual(values = brewer.pal(n=5,"RdYlBu"), breaks = c("Never", "Rarely", "Sometimes", "Often", "Always")) +
  guides(fill = guide_legend(title="Responses", reverse = TRUE))
ggsave("../imgs/diverge_no.png", height = 3, width = 9)
```


## Stripped Down Image
```{r no-text, message=F}
plot(likert(dental_hyg[,c(1:3)], grouping = dental_hyg[,4]), plot.percent.low = FALSE, plot.percent.high = FALSE, plot.percent.neutral = FALSE, legend.position="right") +
  scale_fill_manual(values = brewer.pal(n=5,"RdYlBu"), breaks = c("Never", "Rarely", "Sometimes", "Often", "Always")) +
  guides(fill = guide_legend(title="Responses", reverse = TRUE))+
 theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        # axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
       legend.position = "none")
ggsave("../imgs/likert_card.png", height = 3, width = 3)
```